Article Title

### 2014-01-24 plos.tex

\begin{FPfigure}
    \begin{center}
        \includegraphics[width=17.35cm]{./figures/Figure05.pdf}
    \end{center}
    \caption{{\bf Example of the structure revealed by $C_{\sf sparse+latent}$ the sparse+low-rank estimator.}
    {\bf A and D.} The regularized estimate $C_{\sf sparse+latent}$ closely approximates the sample correlation matrix $C_{\sf 0}$. 
    {\bf B and E.} However, the partial correlation matrices from the two estimates differ substantially.
    {\bf C.} The partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.2\% off-diagonal zeros (bottom-left) and low-rank component of rank 15 (top-right).
    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations: the gray interval indicates the range of correlations containing 82.2\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix. The significant correlations were outside this interval. Yet 58.3\% of the interactions inferred by $C_{\sf sparse+latent}$  linked pairs of neurons whose correlation was below the threshold. Only the remaning 41.7\% of the interactions overlapped with sample correlations above the theshold.
    {ebf G.} A graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons. The line density is proportional to the magnitude of the correlation.
    {\bf H.} A subset of neurons from the center of the cluster shown in {\bf G} showing the regularized partial correlations.
    {\bf I.} The same subset with sample correlations thresholded to match the sparsity of the regularized interactions.
}
\label{fig:05}
\end{FPfigure}

@@

The results of model selection must be interpreted with caution.  As we demonstrated by simulation, a simple model can produce a better estimate even if it does not correctly represent the real dependencies in the data than a more complex model with correct representation of dependencies.   Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the inference that it reveals a better representation of dependencies in the data.  This caveat is related to the phenomenon known as \emph{Stein’s Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.

\paragraph{Physiological interpretation and future directions}
Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.  The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.  The inferred graph of interactions was substantially different from the graph of the most significant sample correlations \figref{05}{F, H, and I}.  The $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that are not present in the sample correlation matrix (\figref{05}{F}).  These negative interactions are produced as explanations for low sample correlation values where high values are expected based on common correlations with other neurons.  The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance.  The resulting network had a larger fraction of negative interactions than in the sample correlation matrix. The negative interactions had a distinct spatial and functional organization from the positive interactions.  To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.

@@

We propose that the structure of the most efficient among many estimators provides a better picture of the functional connectivity than that obtained from unbiased sample correlations.

>>>>

\begin{FPfigure}
    \begin{center}
        \includegraphics[width=17.35cm]{./figures/Figure05.pdf}
    \end{center}
    \caption{{\bf Example of the structure revealed by $C_{\sf sparse+latent}$ the sparse+latent estimator.}
    {\bf A and D.} The regularized estimate $C_{\sf sparse+latent}$ closely approximates the sample correlation matrix $C_{\sf 0}$. 
    {\bf B and E.} However, the partial correlation matrices from the two estimates differ substantially.
    {\bf C.} The partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.2\% off-diagonal zeros (bottom-left) and low-rank component of rank 15 (top-right).
    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations: the gray interval indicates the range of correlations containing 82.2\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix. The significant correlations were outside this interval. Yet 58.3\% of the interactions inferred by $C_{\sf sparse+latent}$  linked pairs of neurons whose correlation was below the threshold. Only the remaning 41.7\% of the interactions overlapped with sample correlations above the theshold.
    {ebf G.} A graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons. The line density is proportional to the magnitude of the correlation.
    {\bf H.} A subset of neurons from the center of the cluster shown in {\bf G} showing the regularized partial correlations.
    {\bf I.} The same subset with sample correlations thresholded to match the sparsity of the regularized interactions.
}
\label{fig:05}
\end{FPfigure}

@@

The results of model selection must be interpreted with caution.  As we demonstrated by simulation, a simple model can produce a better estimate even if it does not correctly represent the real dependencies in the data than a more complex model with correct representation of dependencies.   Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the inference that it reveals a better representation of dependencies in the data.  This caveat is related to the phenomenon known as \emph{Stein's Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.

\paragraph{Physiological interpretation and future directions}
Although the graph of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.  The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.  The inferred graph of interactions differed substantially from the graph of most significant sample correlations \figref{05}{F, H, and I}.  The $C_{\sf sparse+latent}$ estimator revealed a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}).  These negative interactions are often produced as explanations for low sample correlation values where high values were expected based on common correlations with other neurons.  The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance.  The resulting network had a larger fraction of negative interactions than in the sample correlation matrix. The negative interactions had a distinct spatial and functional organization from the positive interactions (\figref{fig:07}).  To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.

The estimator effectively decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.  Since the distinction led to improved estimation, the two types of components could be further investigated.  For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected on the common components/latent units but much less on the pairwise interactions. Previous studies found that correlations measured under anesthesia retained little in common to correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.  When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but does not affect the pairwise interactions as strongly.  The correlation structure also changes as a function of the stimulus condition \cite{Cotton:2013} on time scales that are much faster than changes in the anatomical connectivity. Descriptions of specific changes in the functional connectivity due to stimulus properties could shed light on how circuits orchestrate their activity to represent diverse stimuli.

Here we showed that among several models we tested, we found that for cortical microcircuits, a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix.  This finding valuable in itself: it can be used to improve studies aiming to understand the nature of population coding. Moreover, this estimation approach also provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.  For example, we find a specific structure of negative interactions which may indicate inhibitory circuitry. Future experiment combining molecular markers for different cell types or follow-up multiple patching in slices \cite{Ko:2012}  with our methods can be used to relate functional connectivity to the circuit architecture. Finally, the graphs of interactions derived \emph{in vivo} could potentially reveal how circuits reconfigure dynamically under different brain states or external stimulation: this opens a window to describe what kinds of interactions are activated by stimulation conditions. Extending the estimation framework to temporal domain is important for future work.

Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.  The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.  The inferred graph of interactions was substantially different from the graph of the most significant sample correlations \figref{05}{F, H, and I}.  The $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that are not present in the sample correlation matrix (\figref{05}{F}).  These negative interactions are produced as explanations for low sample correlation values where high values are expected based on common correlations with other neurons.  The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance.  The resulting network had a larger fraction of negative interactions than in the sample correlation matrix. The negative interactions had a distinct spatial and functional organization from the positive interactions.  To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.

@@

\begin{enumerate}
    \item Better estimator of correlations -- better studies of coding, etc.
    \item The interaction model may give more relevant - validation with new experiments 
    \item Hopefully go beyond static connectivity -- but real-time, emergent, connectivity 
    \item The ideal  experiments  would have causal manipulations 
\end{enumerate}
